### 2019-12-15 docs/paper.tex
add model and results sections

\documentclass{article}

\usepackage[final]{neurips_2019}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\graphicspath{ {./images/} }

\title{Music Classification Using Convolutional Neural Networks}

\author{%
  Abel Putovici\And
	Zsombor Molnár
}

\begin{document}

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

\section{Introduction}

Music is and has always been one of the most popular forms of art and 
entertainment. The number of people who use streaming services to consume music 
has grown to a magnitude of millions in this decade. These services make use of 
complicated and complex recommender systems to help grow the number of 
daily active users and to improve their experience of the service. These recommender 
systems rely on extremely large labeled datasets in order to work properly, and 
labeling these datasets manually is a very resource-heavy endeavour.

\subsection{Our Goal}

\paragraph{Music Genres in General}

One possible way of labeling songs is doing it by genre though this introduces 
other difficulties, since these are ment to indicate the artistic nature of music, an 
aspect which tends to be highly subjective and controversial. Music genres do not have a 
proper, formal definition and often overlap as well, making music classification by 
genre a more difficult task to do right.

\paragraph{Our Primary Goal}

Our primary goal was do deliver a solution which can effectively recognize some of the 
most popular and generic music genres and assign these as labels to songs of music.

\section{Our Approach}

There are many existing solutions and research papers regarding the subject and these 
show absolutely usable methods and results, though each of them tend to have a major 
drawback: every solution aims to work in \emph{centralized environments} due to their 
computational requirements. We decided that our approach had to be able to work on 
edge devices such as smart phones or other, more constrained environments, even on 
cost of precision. This decision implied a rather small model size and great model 
simplicity, this is why we chose to try a simple convolutional neural network.

\section{The Model}

\subsection{Trial and Error}

We went through numerous iterations and tried various convolutional neural 
network architectures during development. At first we tried to use convolutional 
windows that would overlap the whole frequency domain and move only in the time domain, 
since although spectrograms can be regarded as arbitrary two dimensional data, the two 
axises have different meanings. This path led nowhere, the model was not able to 
overfit on a small subset of the dataset. We decided to use simple, rectangular convolutional 
windows, which could be called standard by now. This approach was indeed able to overfit, meaning 
it was worth to try learning on the whole dataset. Validation error would cease to decrease 
after the first few epochs. This would be followed by decreasing the input size of the model 
by splitting the samples of 30 seconds length into 4 smaller fragments.

\subsection{The Final Model Architecture}

The final model consists of the following layers:

\begin{itemize}
\item a convolutional layer with a depth of 32 and a window shape of (2, 6)
\item another convolutional layer with a depth of 32 and a window shape of (2, 6)
\item a convolutional layer with a depth of 64 and a window shape identical to the previous ones
\item a fully connected layer of 512 neurons
\end{itemize}

Each convolutional layer was followed by a batch normalization 
layer and a max-pooling layer. L1 and L2 regularizers were adopted to the fully connected layer. 
The model has an output layer of 10 neurons with \emph{softmax} as activation function.

\subsection{Hyperparameters}
\paragraph{Activation functions}
At first we tried using \emph{ReLU} activations between each layer. This lead to many issues, 
since in the context of Mel spectrograms negative numbers have significance. Perceiving this, 
we decided to use \emph{Hyperbolic Tangent} as activation function.

\paragraph{Loss Function}
The loss function used is \emph{Sparse Categorical Crossentropy} since the labels are not one-hot encoded.

\paragraph{Optimizer}
The optimizer algorithm used during training was Stochastic Gradient Descent with a learning rate of 0.01. 
At first we've tried with \emph{Adam} but it did not perform quite as expected. The next choice 
was \emph{Adadelta} which showed promising results, but it did not appear to stable due to the 
fluctuation we have observed on both the graphs of training and validation accuracy.

\paragraph{Number of Epoch and Data Split}
The model was trained for 100 epochs with a batch size of 5. 80 percent of the data was used 
for training, 10 percent was used for validation and the last 10 percent was used for testing.

\paragraph{Vanishing Gradient}

During the training process we've noticed that after a certain number of epochs 
the validation accuracy would cease to converge. This was due to the Vanishing 
Gradient problem, which led to the choice of using \emph{Linear} activation between the 
convolutional layers, leaving \emph{Hyperbolic Tangent} only between the output and 
the fully connected layer.

\section{Results}

\subsection{Training and Validation}

The model is still overfitting with the validation accuracy being stuck at \~60\% and the training accuracy soaring.

\begin{center}
    \includegraphics[scale=0.5]{acc}
		\includegraphics[scale=0.5]{loss}
\end{center}

\subsection{Testing Accuracy}

The final testing accuracy of the model is \textbf{0.61\%}. Below are the Classication Matrix and the final 
Classification Report.

\begin{table}[h]
  \caption{Confusion Matrix}
  \centering
  \begin{tabular}{l | llllllllll}
    Genre & Metal & Classical & Reggae & Jazz & Country & Rock & Hiphop & Disco & Pop & Blues \\
    \midrule
    Metal & 44 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
		Classical & 0 & 28 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0 \\
		Reggae & 1 & 1 & 18 & 0 & 1 & 2 & 5 & 3 & 8 & 5 \\
    Jazz & 1 & 11 & 0 & 28 & 1 & 1 & 0 & 0 & 1 & 2 \\
		Country & 0 & 2 & 0 & 3 & 20 & 1 & 0 & 0 & 0 & 4 \\
		Rock & 12 & 0 & 0 & 2 & 5 & 16 & 4 & 2 & 1 & 4 \\
		Hiphop & 7 & 0 & 1 & 0 & 0 & 0 & 28 & 0 & 1 & 2 \\
		Disco & 4 & 1 & 4 & 0 & 1 & 9 & 7 & 8 & 7 & 2 \\
		Pop & 0 & 3 & 1 & 0 & 3 & 2 & 2 & 0 & 35 & 2 \\
		Blues & 1 & 1 & 0 & 5 & 2 & 0 & 0 & 1 & 0 & 20 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h]
  \caption{Classification Report}
  \centering
  \begin{tabular}{l | l l l l}
    Genre & Precision & Recall & F1-Score & Support (number of samples) \\
    \midrule
       metal   &    1.00   &   0.63  &    0.77    &    70 \\
   classical    &   0.93    &  0.60  &    0.73    &    47 \\
      reggae    &   0.41    &  0.75   &   0.53    &    24 \\
        jazz    &   0.63    &  0.74   &   0.68    &    39 \\
     country    &   0.67    &  0.59   &   0.62    &    34 \\
        rock    &   0.35    &  0.50   &   0.41    &    32 \\
      hiphop    &   0.72    &  0.61   &   0.66    &    46 \\
       disco    &   0.19    &  0.57   &   0.28    &    14 \\
         pop    &   0.73    &  0.66   &   0.69    &    53 \\
       blues    &   0.67   &   0.49   &   0.56   &     41 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h]
  \caption{Classification Report Averages}
  \centering
  \begin{tabular}{l | l l l l}
    Type & Precision & Recall & F1-Score & Support (number of samples) \\
    \midrule
		Macro Average & 0.63 & 0.61 & 0.59 & 400 \\
		Weighted Average & 0.71 & 0.61 & 0.64 & 400 \\
    \bottomrule
  \end{tabular}
\end{table}

\section*{References}

References follow the acknowledgments. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references. {\bf Remember that you can use more than eight
  pages as long as the additional pages contain \emph{only} cited references.}
\medskip

\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.

\end{document}